---
title: "covid 19"
author: "Emma Treat"
date: "10/06/21"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$state<-as.factor(kRdata$state)
kRdata$result<-as.factor(kRdata$result)
kRdata$sex<-as.factor(kRdata$sex)
kRdata$Sex<-as.factor(kRdata$Sex)


```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
```


   
From this bar chart, we can see that more women are alive then men, with women being more prominent then men. The amount of women and men that died was not substantial. All and all, there were more men that died then women.   




```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
We can see from the second bar chart that about 55-56% of women were alive vs about 40% of men, and from the dead catergory we can see more men died vs the amount of women. From both bar charts we can see that women did better in this study. 
    





  
## Numerical results


## Inferential results 




## Conclusions

