---
title: "covid 19"
author: "Emma Treat"
date: "10/06/21"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$state<-as.factor(kRdata$state)
kRdata$result<-as.factor(kRdata$result)
kRdata$sex<-as.factor(kRdata$sex)
kRdata$Sex<-as.factor(kRdata$Sex)


```


### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
```


   
From this bar chart, we can see that more women are alive then men, with women being more prominent then men. The amount of women and men that died was not substantial. All and all, there were more men that died then women.   




```{r}
barchartGC(~result + Sex,data=kRdata, type="percent")

```
  
We can see from the second bar chart that about 55-56% of women were alive vs about 40% of men, and from the dead catergory we can see more men died vs the amount of women. From both bar charts we can see that women did better in this study. 
    





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```



We now have the correct precentages, whereas we can see that the alive group was F 56% and M 44%. In the dead group F ranged in at 36% and M came in at 64%. So we can see here there is a backward relationship between males and females. Females were prodominalty alive wheras men were prodominatley dead. 1.4 F ended up dead, where 3% of males ended up dead. Males deaths were double the rate of women.






## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
```


  
The chi sq test provides the exact numbers. It shows that 1831 were females and males were 1411 that fell in the alive category. 25 women died vs 44 males died. the odds of a female dying is (25/1831). The odds of a male dying is (44/1411). The odds ratio of a male dying to a female is (44/1411)/(25/1831). The odds ratio is 2.3.   






```{r}
fisher.test(table1)
```


The fisher test gves us the odds of dying for males is (44/1411) and the odds of the females dying is (25/1831). When you divide these two numbers you get 2.3. The odds ratio shows that males are 2.3 times more likely to die. We also see the confidence intervals are on the low end 1.4 to 3.9 being the high end. 



## Conclusions


In conclusion we can see a big statistical differnce between men and women. Seemingly more men die of covid vs women. Our p-value is very small, so therefore we reject the null hypothesis. More importantly we can see more men die from this disease vs women.
